ms_figures

### 2018-03-21 man/ms_1_material_methods.Rmd
update

We computed the values of these indices for tree growth (BAI) and greenness during each drought event. We considered 2005 and 2012 as single drought events. The predrought and postdrought values of each target variable (*i.e.*: BAI or EVI) were computed as the mean value during a period of three years before and after the drougth event respectively. A period of three years was chosen because we found similar results comparing periods of two, three and four years (Appendix S7) and this length was used in other studies (e.g. @Gazol2017). In addition we calculated the same indices for BAI data for the most severe drought events (Appendix S4) since 1900.

>>>>

We computed the values of these indices for tree growth (BAI) and greenness during each drought event. We considered 2005 and 2012 as single drought events. The predrought and postdrought values of each target variable (*i.e.*: BAI or EVI) were computed as the mean value during a period of three years before and after the drougth event respectively. A period of three years was chosen because we found similar results comparing periods of two, three and four years (Appendix S6) and this length was used in other studies [*e.g.* @Gazol2017]. In addition we calculated the same indices for BAI data for the most severe drought events (Appendix S3) since 1900.

### 2018-03-21 man/ms_figures.Rmd
update

```{r, fig.height=5}
limits <- as.Date(c("1949-01-01", "2015-12-01"))
spei %>% 
  ggplot(aes(x=date, y=spei6, fill=sign)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada',
    x= '',
    y='SPEI') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", limits=limits) 
```

**Appendix S2**. Drought severity in the Sierra Nevada for the 1950-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). We used data from Global SPEI database. We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada.

@@

##### 
```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% as.data.frame()

@@

**Appendix S4.** Tabla de drought severas.

##### 
**Appendix S5.** Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.

@@

**Appendix S6.** Reilience vs. droughts

##### 
**Appendix S7.** Correlation between indices of resilience (Rt, resistance; Rc, recovery; Rs, Resilience) using periods of several lengths (2, 3 and 4 years after a drought). Top plots (a, b and c) showing the resilience indices of greenness (EVI) to drought; and bottom plots (d, e, f) the resilience indices of tree-growth (BAI) to drought. ws2, ws3 and ws4 indicate periods of 2, 3 and 4 years after a drought.

@@

    df <- df %>% 
    mutate(w1 = case_when(
      w1 == 2 ~ 'ws2',
      w1 == 3 ~ 'ws3',
      w1 == 4 ~ 'ws4'),
      w2 = case_when(
      w2 == 2 ~ 'ws2',
      w2 == 3 ~ 'ws3',
      w2 == 4 ~ 'ws4'))

@@

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, type='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)

>>>>

**Appendix S2**. Drought severity in the Sierra Nevada for the 1950-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). Data from Global SPEI database (http://spei.csic.es/database.html). We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada.

```{r, fig.height=5}
limits <- as.Date(c("1949-01-01", "2015-12-01"))
spei %>% 
  ggplot(aes(x=date, y=spei6, fill=sign)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) +
  theme_bw()  + 
  labs(title= 'Drought evolution for Sierra Nevada',
    x= '',
    y='SPEI (12-month scale)') +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", limits=limits) 
```

@@

##### 
```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% 
  filter(d_duration > 2) %>%  
  mutate(year = ifelse(minyear == maxyear, 
                       minyear, 
                       paste0(minyear, '-', maxyear))) %>% 
  dplyr::select(-minyear, -maxyear, -cat, -month_peak) %>% 
  as.data.frame()

colnames(dys) <- c("Duration (months)", "Intensity", "Severity", "Lowest SPEI", "Year")

@@

**Appendix S3.** Drought events for Sierra Nevada based on SPEI index. A drought event starts in the month when SPEI falls below the threshold of -1.28 [@Pascoa2017]. A drought event is considered only when SPEI value are below threshold for at least two consecutive months [*e.g* @Spinoni2015; Spinoni2017a; @Pascoa2017]. The ***duration*** of a drought event is the number of consecutive months with the SPEI lower than a certain threshold. ***Severity*** of a drought event is the sum of the SPEI values (absolute values) during the duration of the drought event. ***Intensity*** and ***Lowest SPEI** refer to the mean and lowest value of SPEI respectively during the drought event duration.

##### 
**Appendix S4.** Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.

@@

**Appendix S5.** Reilience vs. droughts

##### 
**Appendix S6.** Correlation between indices of resilience (*Rt*, resistance; *Rc*, recovery; *Rs*, Resilience) using periods of several lengths (2, 3 and 4 years after a drought). Top plots (a, b and c) showing the resilience indices of greenness (EVI) to drought; and bottom plots (d, e, f) the resilience indices of tree-growth (BAI) to drought.

@@

    df <- df %>% 
    mutate(w1 = case_when(
      w1 == 2 ~ '2 years',
      w1 == 3 ~ '3 years',
      w1 == 4 ~ '4 years'),
      w2 = case_when(
      w2 == 2 ~ '2 years',
      w2 == 3 ~ '3 years',
      w2 == 4 ~ '4 years'))

@@

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, tpe='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)
